### 2014-11-06 conclusions.tex
More improvements.

We think that our particular approach is applicable in distributed systems as we
have seen. Explore functional programming and coordination?

>>>>

We have presented a novel way of adding coordination to a declarative language without
changing the nature of the language or introducing non-declarative constructs. We
took advantage of the fact that LM uses linear logic, which allows us to derive
coordination facts that can be deleted in order to perform actions in the underlying
runtime system. In the other hand, sensing facts allow the programmer to reason
about locality and scheduling of computation in a data-driven fashion.
We have presented several applications where a judicious use of coordination can lead to
better performance behavior.

We would like to explore this coordination paradigm in distributed systems,
where data locality is far more important than in shared memory systems. We also
think that LM may be useful as a glue distributed/parallel language since it can
easily model any computation pattern using graphs. Furthermore, we think that
other declarative paradigms would benefit from using a similar approach. For
instance, in functional programming, one may consider annotation functions that
pick the right granularity for each problem.

### 2014-11-06 implementation.tex
More improvements.

>>>>

\subsubsection{Coordination overhead}

In the paper~\cite{cruz-ppdp14}, the LM virtual machine was measured and
compared against implementations of similar programs but written in other
languages. For instance, the LBP program was found to be around 2 times slower
than GraphLab, while N Queens program was found to be around 10 to 15 slower
than a sequential C implementation. Some of those programs were compared against
a Python implementations and LM faired fairly well. In this section, we measure
the impact of the coordination mechanisms we have implemented for the new
virtual machine.

Fig.~\ref{results:comparison1} shows the comparison between the original virtual
machine and the new virtual machine with coordination mechanisms. All the
programs benchmarked do not use any kind of coordination. As expected, there
is no degration of performance by adding coordination mechanisms.

\begin{figure}[h!]
   \begin{center}
      \subfloat[]{\includegraphics[width=9cm]{results/comparison1.png}}
   \end{center}
   \caption{Measuring the overhead of coordination mechanisms. The \textbf{Old}
   bars represent the performance of the original virtual machine, while
   \textbf{New} is the relative performance of the new version that includes
   coordination mechanisms.}
   \label{results:comparison1}
\end{figure}

### 2014-11-06 splash-bp.tex
More improvements.

\begin{description}
   \item[Tree building]: Each node has a \texttt{waiting} fact that is used to
   start the tree building process. When the highest priority is picked, a token
   is created that will navigate through the tree. Note that in the rule located
   in lines 14-21, newly added nodes to the tree must have a positive priority
   (due to new belief updates) and must be residing in the same thread.
   We want the tree to be kept in the same thread in order to maximize
   parallelism with 1 tree per thread.
   \item[First phase]: In the third rule (lines 11-12), when the number of nodes
   of the tree reaches a certain limit, a 
   \texttt{first-phase} is generated in order to update the beliefs of all nodes
   in the tree, starting from the leaves and ending at the root
   As the nodes are updated, an \texttt{update} fact is derived to update
   the belief values (line 35).
   \item[Second phase]: In the second phase, the computation of beliefs is
   performed from the root to the leaves and the belief values are updated a
   second time (line 42).
\end{description}

@@

// TREE BUILDING
// continue tree
waiting(A), token(A, All, Next) -o token(A, All, Next).
// start tree
waiting(A), *underline{@priority(A, A, P)}, P > 0.0
   -o token(A, [A], [A]).
// end tree building
token(A, All, Next), length(All) > maxnodes
   -o first-phase(A, All, reverse(All)).
// expand tree
token(A, All, [A | Next])
   -o [collect => L | Side | !edge(A, L, Side),
         0 = count(All, L),
         0 = count(Next, L),
         *underline{@priority(A, L, P)}, P > 0.0,
         *underline{@cpu-id(A, L, Id1)},
         *underline{@cpu-id(A, A, Id2)}, Id1 = Id2 |
         send-token(A, All, Next ++ L)].

send-token(A, All, [])
   -o first-phase(A, All, reverse(All)).
send-token(A, All, [B | Next])
   -o *underline{schedule-next(B)},
      token(B, All ++ [B], [B | Next]).

@@

// SECOND PHASE
second-phase(A, [], _)
   -o *underline{set-priority(A, 0.0)}, waiting(A), update(A).
second-phase(A, [A], Back)
   -o update(A), waiting(Back),
      waiting(A), *underline{set-priority(A, 0.0)}.
second-phase(A, [A, B | Next], Back)
   -o update(A), waiting(Back), *underline{schedule-next(B)},
      second-phase(B, [B | Next], A).
\end{Verbatim}
  \caption{Coordination code for the Splash Belief Propagation program. LM needs
     50 lines of rules to implement splash trees, while GraphLab needs a
     total of 350 lines of C++ to implement the same functionality.  Note
     that when a linear fact is prefixed by \texttt{@}, it indicates that the
     fact is going to be re-derived.}
  \label{code:sbp}
\end{figure}
\normalsize

>>>>

\begin{description}
   \item[Tree building]: Each node has a \texttt{inactive} fact that is used to
   start the tree building process. When the highest priority node is picked, a
   \texttt{tree} is created that will navigate through the tree. In lines 15-21,
   we use an \emph{aggregate}~\cite{cruz-iclp14} to gather all the neighbor
   nodes that have a positive priority (due to a new belief update) and are in the
   same thread. Nodes are collected into list \texttt{L} (line 21) and
   appended to list \texttt{Next}.

   \item[First phase]: In the
   third rule (lines 11-12), when the number of nodes of the tree reaches a
   certain limit, a \texttt{first-phase} is generated to update the beliefs of
   all nodes in the tree, starting from the leaves and ending at the root As the
   nodes are updated, an \texttt{update} fact is derived to update the belief
   values (line 35).

   \item[Second phase]: In the second phase, the computation of beliefs is
   performed from the root to the leaves and the belief values are updated a
   second time (line 42).
\end{description}

@@

// TREE BUILDING
// expand tree by adding neighbor nodes
inactive(A), tree(A, All, Next) -o expand-tree(A, All, Next).
// start tree since we do not have one
inactive(A), *underline{@priority(A, A, P)}, P > 0.0
   -o expand-tree(A, [A], [A]).
// end tree building
expand-tree(A, All, Next), length(All) >= maxnodes
   -o first-phase(A, All, reverse(All)).
// expand tree
expand-tree(A, All, [A | Next]), length([A | Next]) < maxnodes
   -o [collect => L | Side | !edge(A, L, Side),
         0 = count(All, L), // L is not in All
         0 = count(Next, L), // L is not in Next
         *underline{@priority(A, L, P), P > 0.0,}
         *underline{@cpu-id(A, L, Id1)},
         *underline{@cpu-id(A, A, Id2), Id1 = Id2} |
         send-tree(A, All, Next ++ L)].

send-tree(A, All, [])
   -o first-phase(A, All, reverse(All)).
send-tree(A, All, [B | Next])
   -o *underline{schedule-next(B)},
      tree(B, All ++ [B], [B | Next]).

@@

// SECOND PHASE
second-phase(A, [], _)
   -o *underline{set-priority(A, 0.0)}, inactive(A), update(A).
second-phase(A, [A], Back)
   -o update(A), inactive(Back),
      inactive(A), *underline{set-priority(A, 0.0)}.
second-phase(A, [A, B | Next], Back)
   -o update(A), inactive(Back), *underline{schedule-next(B)},
      second-phase(B, [B | Next], A).
\end{Verbatim}
  \caption{Coordination code for the Splash Belief Propagation program. LM needs
     50 lines of rules to implement splash trees, while GraphLab needs a
     total of 350 lines of C++ to implement the same functionality.  Note
     that when a linear fact is prefixed by \texttt{@}, it indicates that the
     fact is going to be re-derived.}
  \label{code:sbp}
\end{figure}
\normalsize
